---
title: "R Notebook"
output:
  html_notebook: default
  html_document:
    df_print: paged
---

```{r}
library(RSQLite)
library(DBI)
```

```{r}
db <- 'chinook.db'

run_query <- function(q) {
  conn <- dbConnect(SQLite(), db)
  result <- dbGetQuery(conn, q)
  dbDisconnect(conn)
  return(result)
}
```

```{r}
show_tables <- function() {
  q = "SELECT name, type FROM sqlite_master WHERE type IN ('table', 'view')"
  return(run_query(q))
}

show_tables()
```
#Business questions
```{r}
albums_to_purchase = '
WITH usa_tracks_sold AS
(
  SELECT il.* FROM invoice_line il
  INNER JOIN invoice i on il.invoice_id = i.invoice_id
  INNER JOIN customer c on i.customer_id = c.customer_id
  WHERE c.country = "USA"
)
SELECT 
  g.name genre, a.name artist, al.title album,
  count(uts.invoice_line_id) tracks_sold,
  cast(count(uts.invoice_line_id) AS FLOAT) / (
    SELECT COUNT(*) from usa_tracks_sold
  ) percentage_sold
FROM usa_tracks_sold uts
INNER JOIN track t on t.track_id = uts.track_id
INNER JOIN genre g on g.genre_id = t.genre_id
LEFT JOIN album al on al.album_id = t.album_id
LEFT JOIN artist a on a.artist_id = al.artist_id
GROUP BY 3
ORDER BY 4 DESC
LIMIT 40;
'
run_query(albums_to_purchase)
```
```{r}
library(ggplot2)
genre_sales <- run_query(albums_to_purchase)
ggplot(data = genre_sales, 
       aes(x = reorder(genre, -percentage_sold), y = percentage_sold)) +
  geom_bar(stat = 'identity')
```
Most sold albums are:
- Nirvana	From The Muddy Banks Of The Wishkah [live]
- Jimi Hendrix	Are You Experienced? 
- The Doors	The Doors
- Marvin Gaye	Seek And Shall Find: More Of The Best (1963-1981)

```{r}
employee_sales_performance = '
WITH customer_support_rep_sales AS
    (
     SELECT
         i.customer_id,
         c.support_rep_id,
         SUM(i.total) total
     FROM invoice i
     INNER JOIN customer c ON i.customer_id = c.customer_id
     GROUP BY 1,2
    )
SELECT
    e.first_name || " " || e.last_name employee,
    e.hire_date,
    SUM(csrs.total) total_sales
FROM customer_support_rep_sales csrs
INNER JOIN employee e ON e.employee_id = csrs.support_rep_id
GROUP BY 1;
'
run_query(employee_sales_performance)
```
```{r}
employee_sales <- run_query(employee_sales_performance)
ggplot(data = employee_sales, 
       aes(x = reorder(employee, -total_sales), y = total_sales)) +
  geom_bar(stat = 'identity')
```
There is no significant difference in sales between employees. There is no data about hours worked / salary, therefore it doesn't really tell anything about productivity.

```{r}
sales_by_country = '
WITH country_or_other AS
    (
     SELECT
       CASE
           WHEN (
                 SELECT count(*)
                 FROM customer
                 where country = c.country
                ) = 1 THEN "Other"
           ELSE c.country
       END AS country,
       c.customer_id,
       il.*
     FROM invoice_line il
     INNER JOIN invoice i ON i.invoice_id = il.invoice_id
     INNER JOIN customer c ON c.customer_id = i.customer_id
    )
SELECT
    country,
    customers,
    total_sales,
    average_order,
    customer_lifetime_value
FROM
    (
    SELECT
        country,
        count(distinct customer_id) customers,
        SUM(unit_price) total_sales,
        SUM(unit_price) / count(distinct customer_id) customer_lifetime_value,
        SUM(unit_price) / count(distinct invoice_id) average_order,
        CASE
            WHEN country = "Other" THEN 1
            ELSE 0
        END AS sort
    FROM country_or_other
    GROUP BY country
    ORDER BY sort ASC, total_sales DESC
    );
'
run_query(sales_by_country)
```

```{r}
library(tidyverse)
country_sales <- run_query(sales_by_country)
country_sales <- country_sales %>% mutate(country = factor(country, levels = country))
bar_plot <- ggplot(data = country_sales, 
       aes(x = country, y = total_sales, fill = country)) +
  geom_bar(stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
plot(bar_plot)
```

```{r}
pie <- ggplot(data = country_sales, 
       aes(x = "", y = total_sales, fill = country)) +
  geom_bar(width = 1, stat = 'identity') + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  coord_polar("y", start=0)
plot(pie)
```

USA is by far the most important country. The other group is large in total too, but exists of 15 countries.

```{r}
albums_vs_tracks = '
WITH invoice_first_track AS
    (
     SELECT
         il.invoice_id invoice_id, il.invoice_line_id, il.unit_price, 
         MIN(il.track_id) first_track_id
     FROM invoice_line il
     GROUP BY 2
    )
SELECT
    COUNT(invoice_line_id) number_of_invoice_lines,
    album_purchase,
    CAST(unit_price AS FLOAT) * (SELECT COUNT(invoice_line_id) AS FLOAT) total_sales
   
FROM
    (
    SELECT
        ifs.*,
        CASE
            WHEN
                 (
                  SELECT t.track_id FROM track t
                  WHERE t.album_id = (
                                      SELECT t2.album_id FROM track t2
                                      WHERE t2.track_id = ifs.first_track_id
                                     ) 
                  EXCEPT 
                  SELECT il2.track_id FROM invoice_line il2
                  WHERE il2.invoice_id = ifs.invoice_id
                 ) IS NULL
             AND
                 (
                  SELECT il2.track_id FROM invoice_line il2
                  WHERE il2.invoice_id = ifs.invoice_id
                  EXCEPT 
                  SELECT t.track_id FROM track t
                  WHERE t.album_id = (
                                      SELECT t2.album_id FROM track t2
                                      WHERE t2.track_id = ifs.first_track_id
                                     ) 
                 ) IS NULL
             THEN "yes"
             ELSE "no"
         END AS "album_purchase"
     FROM invoice_first_track ifs
    )
GROUP BY album_purchase;
'
run_query(albums_vs_tracks)
```

album sales accounts for 30% of total sales. Would not stop buying full albums.


```{r}
Rock = '
SELECT * FROM invoice_line
LIMIT 100
'

run_query(Rock)
```


